---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# magutils

<!-- badges: start -->
[![R-CMD-check](https://github.com/f-hafner/magutils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/f-hafner/magutils/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of magutils is to facilitate loading and extracting data from a database with records from Microsoft Academic Graph and ProQuest Dissertations. **Work in progress.**

## Installation

You can install the development version of magutils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("f-hafner/magutils")
```

## Example

If you do not have access to the full database, use the example database like this:

```{r example}
library(magutils)

db_file <- system.file("extdata", "AcademicGraph.sqlite", package = "magutils")
conn <- connect_to_db(db_file)

```

Then query the graduate links:

```{r}

links <- get_graduate_links(conn)

```






