---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# magutils

<!-- badges: start -->
[![R-CMD-check](https://github.com/f-hafner/magutils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/f-hafner/magutils/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of magutils is to facilitate loading and extracting data from a database with records from Microsoft Academic Graph and ProQuest Dissertations. **Work in progress.**

## Installation

You can install the development version of magutils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("f-hafner/magutils")
```

## Example

If you do not have access to the full database, use the example database like this:

```{r example}
library(magutils)

db_file <- db_example("AcademicGraph.sqlite")
conn <- connect_to_db(db_file)

```

Then query the graduate links:

```{r}
links <- get_links(conn, from = "graduates", lazy = TRUE)
```

Or query info on graduates:

```{r}
graduates <- authors_proquest(conn, lazy = FALSE, limit = 3)
```

You can join the two together

```{r}
library(magrittr)
links <- get_links(conn, from = "graduates", lazy = TRUE)
d_full <- authors_proquest(conn, limit = 5) %>%
  dplyr::left_join(links, by = "goid") %>%
  dplyr::collect()
```


At the end, do not forget to disconnect from the database:

```{r}
DBI::dbDisconnect(conn)
```


