---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# magutils

<!-- badges: start -->
[![R-CMD-check](https://github.com/f-hafner/magutils/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/f-hafner/magutils/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of magutils is to facilitate loading and extracting data from a database with records from Microsoft Academic Graph and ProQuest Dissertations and make the functions available to co-authors and RAs. In the future, we may publish a "back-end" package to generate the database. 

## Installation

You can install the development version of magutils from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("f-hafner/magutils")
```

## Example

If you do not have access to the full database, use the example database like this:

```{r example}
library(magutils)

db_file <- db_example("AcademicGraph.sqlite")
conn <- connect_to_db(db_file)

```

Then query the graduate links:

```{r}
links <- get_links(conn, from = "graduates", lazy = TRUE)
```

Or query info on graduates:

```{r}
graduates <- get_proquest(conn, from = "graduates", lazy = FALSE, limit = 3)
```

You can join the two together

```{r}
library(magrittr)
links <- get_links(conn, from = "graduates", lazy = TRUE)
d_full <- get_proquest(conn, from = "graduates", limit = 5) %>%
  dplyr::left_join(links, by = "goid") %>%
  dplyr::collect()
```


At the end, do not forget to disconnect from the database:

```{r}
DBI::dbDisconnect(conn)
```


## Main functions 


Extracting key tables

- `get_proquest`: Source data on dissertations in United States from ProQuest.

- `get_links`: Load links between ProQuest and MAG. Can be links from PhD graduates to MAG authors, or from PhD advisors to MAG authors


The following functions can be used to get more information from the records in the tables above:

- `graduate_fields`: table with main field of PhD graduates. this is at the unit level

- `augment_tbl`: augment a table with various additional information:

  - output
  
  - affiliations 
  
  - co-authors

- `define_gender`: define the gender of the 



