---
title: "vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(magutils)
```


The package operates on a relational database of bibliometric data from Microsoft Academic Graph and dissertation data from ProQuest. To illustrate the usage, the package ships with a small replicate of the database with the necessary tables. Let's load the database:

```{r}

db_file <- db_example("AcademicGraph.sqlite")
conn <- connect_to_db(db_file)


```
The `tbls` output above shows the all tables in the database.

# Database structure

The database contains both some of the tables originally in MAG and custom tables created by us. The original MAG table schema can be found [here](https://docs.microsoft.com/en-us/academic-services/graph/reference-data-schema#paper-fields-of-study).


## `current_links`
Contains the links between MAG and ProQuest. It is currently created outside of the package. The table has the following columns (columns irrelevant for the package are omitted).

- AuthorId: The author identifier in MAG.
- goid: The dissertation identifier in ProQuest.
- link_score: The probability $\in [0, 1]$ that the linked entities are truely the same. Calculated from the linking algorithm not part of the package.
- iteration_id: The iteration identifer of the linking algorithm; not relevant for this package.


## `FieldsOfStudy`
 
The fields of study as classified by MAG. The relevant columns are 
 
- FieldOfStudyId: field identifier.
- NormalizedName: the name of the field.

## `FirstNamesGender`

Maps first names to gender; data from genderize.io. Columns:

- FirstName: string of first name
- ProbabilityFemale: probability that a person with `FirstName` is female.


## `pq_authors`

Table with authors of PhD dissertations from ProQuest. The relevant columns are:

- goid 
- fullname, firstname
- degree_year: year of PhD degree.
- degree_level: level of the degree; usually PhD or similar.
- university_id: identifier of the university.
- thesistitle: title of the thesis, normalized to match the paper titles in MAG.

## `pq_fields_mag`

The fields of the dissertation and their corresponding field (level 0) in MAG.

- goid
- position: position, starting at 0, as reported in the xml files of the dissertation. The ordering has, to our knowledge, no meaning; but some dissertations have only one field.
- fieldname: the field name as reported in ProQuest
- mag_field0: the corresponding field0 in MAG. We make this correspondence based on similarity of field names and on some tables of fields from the OECD.

## `pq_unis`

Universities in ProQuest

- university_id: university identifer
- normalizedname: name of the university
- location: location of the university. Not standardized, but location of U.S. universities have all a "United States" string in them.



We can use the following query to see the specific contents of each table: 

*TODO: use the simple function here when written*

```{r}

res <- DBI::dbSendQuery(
  conn = conn,
  statement = "select sql from sqlite_master where tbl_name = 'current_links' and type = 'table' "
)
DBI::dbFetch(res)
DBI::dbClearResult(res)

```

# Using the package


## 



Extracting key tables

- `get_proquest`: Source data on dissertations in United States from ProQuest.

- `get_links`: Load links between ProQuest and MAG. Can be links from PhD graduates to MAG authors, or from PhD advisors to MAG authors

- `define_field`: define the field of study for records in a table. 

- `define_gender`: define gender of a table of persons with firstnames.

- `augment_tbl`: augment a table with various additional information: output, affiliations, co-authors. Because `output` and `affiliations` are at the unit-year level, the result will be a table at the unit-year level. I am not sure if this is the best way to do it (also the naming wrt to the previous functions), but we have to see how it works in practice. 
  

```{r}
DBI::dbDisconnect(conn)
```














