---
title: "Vignette magutils"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette magutils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(magutils)
library(magrittr)
```


The package operates on a relational database of bibliometric data from Microsoft Academic Graph and dissertation data from ProQuest. To illustrate the usage, the package ships with a small replicate of the database with the necessary tables. Let's load the database:

```{r}

# TODO: change to small db replicate!

# db_file <- db_example("AcademicGraph.sqlite")
db_file <- "/mnt/ssd/AcademicGraph/AcademicGraph.sqlite"
conn <- connect_to_db(db_file)


```
The `tbls` output above shows the all tables in the database.

# Introducing the database

## Overview

The database contains some of the tables originally in MAG and custom tables created by us. The tables taken from MAG have the same name as in MAG; the schema to their database is [here](https://docs.microsoft.com/en-us/academic-services/graph/reference-data-schema#paper-fields-of-study).

**Naming conventions**

The names of the tables in the database and the columns in the tables follow two different conventions.
Tables from MAG are named with CamelCase, as in the original database; custom tables created by us are named with snake_case (with a few exceptions not relevant to the front-end).
Columns in MAG tables are named with CamelCase; columns in custom tables can be named either with CamelCase or snake_case.

Tables from ProQuest data are indicated with a "pq_" prefix; they and their columns are named with snake_case. 




## Database structure
The following describes each table; columns irrelevant to the package are omitted.


### `current_links` and `current_links_advisors`

Tables with links between records in MAG and ProQuest. They are created outside of the package. 

`current_links` has links between the authors dissertations and authors in MAG. `current_links_advisors` has links between the advisors of dissertations and authors in MAG.


- AuthorId: The author identifier in MAG.
- link_score: The probability $\in [0, 1]$ that the linked entities are truely the same. Calculated from the linking algorithm not part of the package.
- iteration_id: The iteration identifer of the linking algorithm; not relevant for this package.

The identifier in ProQuest differs:

- for `current_links`, it is goid: The dissertation identifier in ProQuest.
- for `current_links_advisors`, it is relationship_id: the combination of dissertation identifier and position of the reported advisor.


### `FieldsOfStudy`
 
The fields of study as classified by MAG. The relevant columns are 
 
- FieldOfStudyId: field identifier.
- NormalizedName: the name of the field.

### `FirstNamesGender`

Maps first names to gender; data from genderize.io. Columns:

- FirstName: string of first name
- ProbabilityFemale: probability that a person with `FirstName` is female.


### `pq_authors`

Table with authors of PhD dissertations from ProQuest. The relevant columns are:

- goid 
- fullname, firstname
- degree_year: year of PhD degree.
- degree_level: level of the degree; usually PhD or similar.
- university_id: identifier of the university.
- thesistitle: title of the thesis, normalized to match the paper titles in MAG.

### `pq_fields_mag`

The fields of the dissertation and their corresponding field (level 0) in MAG.

- goid
- position: position, starting at 0, as reported in the xml files of the dissertation. The ordering has, to our knowledge, no meaning; but some dissertations have only one field.
- fieldname: the field name as reported in ProQuest
- mag_field0: the corresponding field0 in MAG. We make this correspondence based on similarity of field names and on some tables of fields from the OECD.

### `pq_unis`

Universities in ProQuest

- university_id: university identifier
- normalizedname: name of the university
- location: location of the university. Not standardized, but location of U.S. universities have all a "United States" string in them.

### `pq_advisors`

The dissertation advisors.

- goid
- position: advisor's position, starting at 0, as reported in the xml files of the dissertation. The ordering has, to our knowledge, no meaning; but some dissertations have only one field. A position $> 0$ indicates the dissertation reports more than one advisor.
- lastname, firstname: reported names of the advisor
- relationship_id: unique identifier of the student-advisor relationship. This is just the combined `goid` and the `position`. 

Note that advisors do not have their own unique identifier.


## Getting information on the database

We can query information from the database schema with some custom functions.

We can use the following query to see the specific contents of each table: 

*TODO: use the simple function here when written*

```{r}

res <- DBI::dbSendQuery(
  conn = conn,
  statement = "select sql from sqlite_master where tbl_name = 'current_links' and type = 'table' "
)
DBI::dbFetch(res)
DBI::dbClearResult(res)

```

# Using the package


## Loading data from proquest: `get_proquest`

The function loads the dissertations written in the United States (available in ProQuest) between `start_year` and `end_year`. 
The underlying table used for the output depends on `from` and is thus either `pq_authors` when querying "graduates" or `pq_advisors` when querying "advisors".
The final output of the function also depends on `from`, and in general the returned columns always refer to the respective units.

Let's see an example:


```{r}

graduates <- get_proquest(conn, from = "graduates", lazy = FALSE, limit = 3)
head(graduates)

```

The output consists of the following:

- The field of study, using `define_field`: The MAG field 0, mapped from the reported first field of the dissertation. 
- The gender, using `define_gender`: The imputed gender given the first name of the person.

Alternatively, we can query advisors:

```{r}

advisors <- get_proquest(conn, from = "advisors") #TODO: add pq_advisors table for this function
head(advisors)

```

Important here is that the gender refers to the advisor. If we wanted to study student-advisor matches, we could join the output from "graduates" to the output from "advisors", using the column `goid` for the join. Note that a student can have multiple advisors, as indicated by the `position` column.



## Loading the links: `get_links`

This returns the links between records in MAG and in ProQuest. With `from = "graduates"`, we get the links from `current_links`. With `from = "advisors"`, we get the links from `current_links_advisors`. 


```{r}
links <- get_links(conn, from = "graduates", lazy = TRUE)
head(links)
```


```{r}

links <- get_links(conn, from = "advisors", lazy = TRUE)
head(links)

```

As stated in the message, we suspect the links between advisors need a higher `min_score` constraint to omit many false positives.



## Adding information to records: `augment_tbl` 

The function has two purposes:

1. Join output or affiliation information to author units 

2. Join information on affiliations of co-authors of author units


### Join output or affiliation information to author units 

The tables used to augment `tbl` contain multiple rows per AuthorId, as the following example illustrates:


```{r}

graduates_affiliation <- get_links(conn = conn,
                            from = "graduates") %>%
  augment_tbl(conn = conn,
              with = "affiliation") 

head(graduates_affiliation)

```

For each `AuthorId`, we have one row for each year in which there is at least one paper with a reported affiliation of the author. 
The same when using `with = "output"`. 
The function checks if there is already a `Year` column in the `tbl`, and if so, it automatically joins on `AuthorId` and `Year` and therefore avoids duplicates.



### Join information on affiliations of co-authors of author units.

TBD.



## Processing after querying

When should you load the data into memory with `collect()`? I suggest to do as many operations as possible on the database with `dbplyr`. That said, when data from ProQuest and MAG are combined, the resulting tables are usually not very large and should fit into memory.






```{r}
DBI::dbDisconnect(conn)
```














